---
title: "Steak Cutting Analysis"
format: html
editor: visual
embed-resources: true
---

# Set-up

```{r}
#| message: false
library(readxl)
library(lme4)
library(lmerTest)
library(emmeans)
library(janitor)
library(multcomp)
library(multcompView) 
library(tidyverse)
library(sjstats)
```

## Data

```{r}
# load in data

# For Master Set including a, b, l, tep and ph 
color <- read_excel('Data/Meat_Data.xlsx', sheet = 'sorted beef color ')
tbars <- read_excel('Data/Meat_Data.xlsx', sheet = 'sorted beef TBARS')
ph <- read_excel('Data/Meat_Data.xlsx', sheet = 'pH')

# For % yield, microbial colonies and ecoli 
full <- read_excel('Data/Meat_Data.xlsx', sheet = 'everything one sheet')
apc <- read_excel('Data/APC_Data.xlsx')
ecoli <- read_excel('Data/Coliforms_Data.xlsx')
```

```{r}
# clean and combine data
master <- color %>%
  mutate(batch = toupper(batch),
         steak = ifelse(batch == "A" & trt == "KK" & display == 0 & index == 2, "b", steak)) %>%
  left_join(tbars, by = c("batch" = "Batch", "steak" = "patty replicates", "display" = "Display",
                           "index", "trt", "muscle")) %>%
  inner_join(ph, by = c("trt" = "Cutting", "steak", "muscle", "index", "batch", "display")) %>%
  select(batch, trt, display, index, steak, muscle, `average L`, `average a`, `average b`, `calculated TEP equivalent`, `pH-ave`) %>%
  clean_names() %>%
  rename('ph_average' = 'p_h_ave', 'tep' = 'calculated_tep_equivalent') %>%
  mutate(display = as.factor(display))
```

```{r}
# Obtaining bacteria information 
bact <- apc %>%
  inner_join(ecoli) %>%
  mutate(trt = str_sub(type, 1, 2),
         batch = str_sub(type, 3, 3),
         display = as.factor(display)) %>%
  arrange(trt, display) %>%
  select(acp, coliforms)
```

```{r}
# % yield 
full <- full %>% 
  mutate(display_yield_beef = ifelse(display_yield_beef == "*", NA, display_yield_beef),
         display_yield_beef = as.numeric(display_yield_beef),
         display = as.factor(display)) %>%
  inner_join(master, by = c("beef_loin_lightness" = "average_l", "beef_loin_yellowness" = "average_b")) %>% #getting batch col 
  select(trt.x, display.x, index.x, batch, display_yield_beef) %>%
  rename("trt" = "trt.x", 'display' = 'display.x', 'index' = 'index.x', 'yield' = 'display_yield_beef') %>%
  cbind(bact)
  
```

# Linear Mixed Effects Regression

## Functions

```{r}
fit_lmer <- function(rv, alpha = 0.95) {
  
  # Model 
  mod <- lmer(get(rv) ~ trt*muscle*display + (1 | batch), data = master)
  print("Summary")
  print(summary(mod))
  
  print(performance::icc(mod))

  # Contrasts
  ests <- emmeans(mod, specs = ~ trt + display | muscle, infer = c(TRUE, TRUE), level = alpha)
  print("Contrasts")
  print(contrast(ests, method = "pairwise"))

  # Letters 
  lets <- cld(ests, alpha = alpha, Letters = letters)
  print("Letters")
  print(lets)
  
  # Plot 
  # plot <- ests %>% 
  #   as_tibble() %>%
  #   ggplot(aes(x = display, y = emmean, color = muscle)) +
  #     geom_point() +
  #     geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
  #     facet_wrap(~ trt) +
  #     labs(title = paste("Estimated Marginal Means for", rv),
  #          x = "Display",
  #          y = rv) +
  #     theme_bw() +
  #     theme(
  #       panel.grid.minor.x = element_blank(),
  #       panel.grid.minor.y = element_blank()
  #       ) 
  # print(plot)

}

fit_lmer2 <- function(rv, alpha = 0.95) {
  
  # Model 
  mod <- lmer(get(rv) ~ trt*display + (1 | batch), data = full)
  print("Summary")
  print(summary(mod))
  
  print(paste("ICC: ", icc(mod)))

  # Contrasts
  ests <- emmeans(mod, specs = ~ trt + display, infer = c(TRUE, TRUE), level = alpha)
  print("Contrasts")
  print(contrast(ests, method = "pairwise"))

  # Letters 
  lets <- cld(ests, alpha = alpha, Letters = letters)
  print("Letters")
  print(lets)

}
```

## Average L\*

```{r}
fit_lmer("average_l")
```

## Average a

```{r}
fit_lmer("average_a")
```

## Average b

```{r}
fit_lmer("average_b")
```

## TEP

```{r}
fit_lmer("tep")
```

## ph Average

```{r}
fit_lmer("ph_average")
```

## % Yield

```{r}
fit_lmer2("yield")
```

## ACP

```{r}
fit_lmer2("acp")
```

## coliforms

```{r}
fit_lmer2("coliforms")
```

# Interpretations

-   Adjusted ICC: $$ð›”^2_{between}/(ð›”^2_{between} + ð›”^2_{within})$$

    -   Adjusted ICC: Focuses only on random effects, ignoring fixed effects. Useful for analyzing the impact of random effects alone.

    -   Unadjusted ICC: Includes both random and fixed effects in the variability calculation. Useful for a broader understanding of variability, including the influence of fixed effects.

    -   \_\_% of the variability in {response variable} is accounted for by the difference between {grouping variable}.

    -   i.e. 42% of the variability in Average L\* is accounted for by the difference between batches (and the remaining 58% is due to the difference within batches).

    -   NOTE: We are using the Adjusted ICC.

-   ANOVA Output:

    -   trt:
    -   muscle:
    -   display:

-   Letter Plot:
